Experiment 3 of 4

### 2016-09-27 experiments/e3/writeup/e3.Rmd
e3 writeup less verbose

---
title: "Experiment 3 of 4"
output:
  html_document:
    toc: true
    <!--toc_float: true-->
    mathjax: default
    template: default
    code_folding: hide
    theme: default
---  
# Experiment 3

```{r docOpts}
library(knitr)
```

```{r loadLibraries}
suppressPackageStartupMessages(library(MASS))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(lme4))
suppressPackageStartupMessages(library(xtable))
suppressPackageStartupMessages(library(plyr))
suppressPackageStartupMessages(library(reshape))
suppressPackageStartupMessages(library(lattice))
suppressPackageStartupMessages(library(LMERConvenienceFunctions))
suppressPackageStartupMessages(library(car))
suppressPackageStartupMessages(library(grid))
suppressPackageStartupMessages(library(lmerTest))
suppressPackageStartupMessages(library(reshape2))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(require(languageR))
```

# Experiment Structure
Each subject saw 192 trials, being 3 cycles of 64 trials. There were 38 participants. There were 7296 rows in the full data set.

# Trial Procedure
The procedure for each trial was as follows. First an instruction that was a referring expression was presented until participants dismissed it with a key press. Then the squares with dots were presented until the participant responded with a key press, with a 60 second timeout if they did not respond.

# Prepare the data
Source additional R files with functions for computing confidence intervals.

```{r sourceFunctions}
source("functions/CIfunctions.R")
```

Get data.

```{r readdata}
dat <- read.delim("data/allresults.txt", stringsAsFactors = FALSE)

# declare local variables
number_of_valid_subjects <- 38
number_of_rows <- 7296
number_of_trials_per_subject <- number_of_rows / number_of_valid_subjects # 192

# sort out borderline responses into expected near and far
# How many squares in the square they chose? 
dat$RESPONSE <- as.character(dat$RESPONSE)
for (row in 1 : nrow(dat) ) {
  switch(dat[row,'RESPONSE'],
         'LEFT' = {dat[row, 'choice'] <- dat[row, 'Left']},
         'MIDDLE' = {dat[row, 'choice'] <- dat[row, 'Mid']},
         'RIGHT' = {dat[row, 'choice'] <- dat[row, 'Right']}         
  )
}
dat$crossed=paste('Con',dat$Condition,':Quan',dat$Quantity,':Item',dat$Item,sep="")

dat[dat$crossed=='Con1:Quan1:Item1', 'ResponseExpected'] <-  6 
dat[dat$crossed=='Con1:Quan1:Item1', 'ResponseNear']     <- 15  
dat[dat$crossed=='Con1:Quan1:Item1', 'ResponseFar']      <- 24

dat[dat$crossed=='Con1:Quan1:Item2', 'ResponseExpected'] <- 16
dat[dat$crossed=='Con1:Quan1:Item2', 'ResponseNear']     <- 25  
dat[dat$crossed=='Con1:Quan1:Item2', 'ResponseFar']      <- 34

dat[dat$crossed=='Con1:Quan1:Item3', 'ResponseExpected'] <- 26
dat[dat$crossed=='Con1:Quan1:Item3', 'ResponseNear']     <- 35
dat[dat$crossed=='Con1:Quan1:Item3', 'ResponseFar']      <- 44

dat[dat$crossed=='Con1:Quan1:Item4', 'ResponseExpected'] <- 36
dat[dat$crossed=='Con1:Quan1:Item4', 'ResponseNear']     <- 45
dat[dat$crossed=='Con1:Quan1:Item4', 'ResponseFar']      <- 54

dat[dat$crossed=='Con1:Quan2:Item1', 'ResponseExpected'] <- 24
dat[dat$crossed=='Con1:Quan2:Item1', 'ResponseNear']     <- 15 
dat[dat$crossed=='Con1:Quan2:Item1', 'ResponseFar']      <-  6

dat[dat$crossed=='Con1:Quan2:Item2', 'ResponseExpected'] <- 34
dat[dat$crossed=='Con1:Quan2:Item2', 'ResponseNear']     <- 25  
dat[dat$crossed=='Con1:Quan2:Item2', 'ResponseFar']      <- 16

dat[dat$crossed=='Con1:Quan2:Item3', 'ResponseExpected'] <- 44
dat[dat$crossed=='Con1:Quan2:Item3', 'ResponseNear']     <- 35
dat[dat$crossed=='Con1:Quan2:Item3', 'ResponseFar']      <- 26

dat[dat$crossed=='Con1:Quan2:Item4', 'ResponseExpected'] <- 54
dat[dat$crossed=='Con1:Quan2:Item4', 'ResponseNear']     <- 45
dat[dat$crossed=='Con1:Quan2:Item4', 'ResponseFar']      <- 36

dat[dat$crossed=='Con2:Quan1:Item1', 'ResponseExpected'] <-  6 
dat[dat$crossed=='Con2:Quan1:Item1', 'ResponseNear']     <- 15  
dat[dat$crossed=='Con2:Quan1:Item1', 'ResponseFar']      <- 24

dat[dat$crossed=='Con2:Quan1:Item2', 'ResponseExpected'] <- 16
dat[dat$crossed=='Con2:Quan1:Item2', 'ResponseNear']     <- 25  
dat[dat$crossed=='Con2:Quan1:Item2', 'ResponseFar']      <- 34

dat[dat$crossed=='Con2:Quan1:Item3', 'ResponseExpected'] <- 26
dat[dat$crossed=='Con2:Quan1:Item3', 'ResponseNear']     <- 35
dat[dat$crossed=='Con2:Quan1:Item3', 'ResponseFar']      <- 44

dat[dat$crossed=='Con2:Quan1:Item4', 'ResponseExpected'] <- 36
dat[dat$crossed=='Con2:Quan1:Item4', 'ResponseNear']     <- 45
dat[dat$crossed=='Con2:Quan1:Item4', 'ResponseFar']      <- 54

dat[dat$crossed=='Con2:Quan2:Item1', 'ResponseExpected'] <- 24
dat[dat$crossed=='Con2:Quan2:Item1', 'ResponseNear']     <- 15 
dat[dat$crossed=='Con2:Quan2:Item1', 'ResponseFar']      <-  6

dat[dat$crossed=='Con2:Quan2:Item2', 'ResponseExpected'] <- 34
dat[dat$crossed=='Con2:Quan2:Item2', 'ResponseNear']     <- 25  
dat[dat$crossed=='Con2:Quan2:Item2', 'ResponseFar']      <- 16

dat[dat$crossed=='Con2:Quan2:Item3', 'ResponseExpected'] <- 44
dat[dat$crossed=='Con2:Quan2:Item3', 'ResponseNear']     <- 35
dat[dat$crossed=='Con2:Quan2:Item3', 'ResponseFar']      <- 26

dat[dat$crossed=='Con2:Quan2:Item4', 'ResponseExpected'] <- 54
dat[dat$crossed=='Con2:Quan2:Item4', 'ResponseNear']     <- 45
dat[dat$crossed=='Con2:Quan2:Item4', 'ResponseFar']      <- 36

dat[dat$crossed=='Con3:Quan1:Item1', 'ResponseExpected'] <-  6 
dat[dat$crossed=='Con3:Quan1:Item1', 'ResponseNear']     <- 15  
dat[dat$crossed=='Con3:Quan1:Item1', 'ResponseFar']      <- 24

dat[dat$crossed=='Con3:Quan1:Item2', 'ResponseExpected'] <- 16
dat[dat$crossed=='Con3:Quan1:Item2', 'ResponseNear']     <- 25  
dat[dat$crossed=='Con3:Quan1:Item2', 'ResponseFar']      <- 34

dat[dat$crossed=='Con3:Quan1:Item3', 'ResponseExpected'] <- 26
dat[dat$crossed=='Con3:Quan1:Item3', 'ResponseNear']     <- 35
dat[dat$crossed=='Con3:Quan1:Item3', 'ResponseFar']      <- 44

dat[dat$crossed=='Con3:Quan1:Item4', 'ResponseExpected'] <- 36
dat[dat$crossed=='Con3:Quan1:Item4', 'ResponseNear']     <- 45
dat[dat$crossed=='Con3:Quan1:Item4', 'ResponseFar']      <- 54

dat[dat$crossed=='Con3:Quan2:Item1', 'ResponseExpected'] <- 24
dat[dat$crossed=='Con3:Quan2:Item1', 'ResponseNear']     <- 15 
dat[dat$crossed=='Con3:Quan2:Item1', 'ResponseFar']      <-  6

dat[dat$crossed=='Con3:Quan2:Item2', 'ResponseExpected'] <- 34
dat[dat$crossed=='Con3:Quan2:Item2', 'ResponseNear']     <- 25  
dat[dat$crossed=='Con3:Quan2:Item2', 'ResponseFar']      <- 16

dat[dat$crossed=='Con3:Quan2:Item3', 'ResponseExpected'] <- 44
dat[dat$crossed=='Con3:Quan2:Item3', 'ResponseNear']     <- 35
dat[dat$crossed=='Con3:Quan2:Item3', 'ResponseFar']      <- 26

dat[dat$crossed=='Con3:Quan2:Item4', 'ResponseExpected'] <- 54
dat[dat$crossed=='Con3:Quan2:Item4', 'ResponseNear']     <- 45
dat[dat$crossed=='Con3:Quan2:Item4', 'ResponseFar']      <- 36

dat[dat$crossed=='Con4:Quan1:Item1', 'ResponseExpected'] <-  6 
dat[dat$crossed=='Con4:Quan1:Item1', 'ResponseNear']     <- 15  
dat[dat$crossed=='Con4:Quan1:Item1', 'ResponseFar']      <- 24

dat[dat$crossed=='Con4:Quan1:Item2', 'ResponseExpected'] <- 16
dat[dat$crossed=='Con4:Quan1:Item2', 'ResponseNear']     <- 25  
dat[dat$crossed=='Con4:Quan1:Item2', 'ResponseFar']      <- 34

dat[dat$crossed=='Con4:Quan1:Item3', 'ResponseExpected'] <- 26
dat[dat$crossed=='Con4:Quan1:Item3', 'ResponseNear']     <- 35
dat[dat$crossed=='Con4:Quan1:Item3', 'ResponseFar']      <- 44

dat[dat$crossed=='Con4:Quan1:Item4', 'ResponseExpected'] <- 36
dat[dat$crossed=='Con4:Quan1:Item4', 'ResponseNear']     <- 45
dat[dat$crossed=='Con4:Quan1:Item4', 'ResponseFar']      <- 54

dat[dat$crossed=='Con4:Quan2:Item1', 'ResponseExpected'] <- 24
dat[dat$crossed=='Con4:Quan2:Item1', 'ResponseNear']     <- 15 
dat[dat$crossed=='Con4:Quan2:Item1', 'ResponseFar']      <-  6

dat[dat$crossed=='Con4:Quan2:Item2', 'ResponseExpected'] <- 34
dat[dat$crossed=='Con4:Quan2:Item2', 'ResponseNear']     <- 25  
dat[dat$crossed=='Con4:Quan2:Item2', 'ResponseFar']      <- 16

dat[dat$crossed=='Con4:Quan2:Item3', 'ResponseExpected'] <- 44
dat[dat$crossed=='Con4:Quan2:Item3', 'ResponseNear']     <- 35
dat[dat$crossed=='Con4:Quan2:Item3', 'ResponseFar']      <- 26

dat[dat$crossed=='Con4:Quan2:Item4', 'ResponseExpected'] <- 54
dat[dat$crossed=='Con4:Quan2:Item4', 'ResponseNear']     <- 45
dat[dat$crossed=='Con4:Quan2:Item4', 'ResponseFar']      <- 36

dat$isResponseExpected <- dat$choice == dat$ResponseExpected
dat$isResponseNear <- dat$choice == dat$ResponseNear
dat$isResponseFar <- dat$choice == dat$ResponseFar

for ( row in 1:nrow(dat) ) {
  dat[row, 'switchResponse'] <- 
    ifelse(dat[row, 'isResponseExpected']==TRUE, 'Expected', 
           ifelse(dat[row, 'isResponseNear']==TRUE, 'Near', 'Far') ) 
}

## treat variables

# SUBJECT
# ensure Subject is a factor 
dat$Subject=factor(paste("s",sprintf("%02d",dat$Subject),sep=""))

# TRIAL
# Trial for subject, 1 to 192
dat$Trial = rep(x = 1:number_of_trials_per_subject, times = number_of_valid_subjects)

# make a centred Trial for modeling
dat$c_Trl <-dat$Trial - mean(dat$Trial)

# make a scaled Trial for modelling
dat$s_Trl <- as.numeric(scale(dat$Trial))

# ID
# id is a unique identifier for the 7680 row data
dat$id <- factor(paste(paste(dat$Subject), 
                       paste("t", sprintf("%03d", dat$Trial), sep="") , sep=":"))

# ITEM
# create a centred numeric item variable for modeling
dat$c_Itm <- ifelse(dat$Item==1, -.75, 
                    ifelse(dat$Item==2, -.25,
                           ifelse(dat$Item==3, .25, .75)))

# make Item be a factor and assign labels
dat$Item <- factor(dat$Item, levels=c(1,2,3,4), labels=c("06:15:24", "16:25:34", "26:35:44", "36:45:54"))

# add ratios for Item
item_range_ratio = c(6 / 24, 16 / 34, 26 / 44, 36 / 54) 
# 0.2500000 0.4705882 0.5909091 0.6666667
item_range_ratio_scaled = as.vector(scale(c(6 / 24, 16 / 34, 26 / 44, 36 / 54)))
# -1.3441995 -0.1316642  0.5297187  0.9461450
item_mean_ratio = c(mean(c(6 / 15, 15 / 24)), mean(c(16 / 25, 25 / 34)), mean(c(26 /35, 35 / 44)), mean(c(36 / 45, 45 / 54))) 
# 0.5125000 0.6876471 0.7691558 0.8166667
item_mean_ratio_scaled = as.vector(scale(c(mean(c(6 / 15, 15 / 24)), mean(c(16 / 25, 25 / 34)), mean(c(26 /35, 35 / 44)), mean(c(36 / 45, 45 / 54))))) 
# -1.37582241 -0.06614191  0.54334858 0.89861574

dat[dat$Item == "06:15:24", "item_range_ratio"] <-  0.2500000
dat[dat$Item == "16:25:34", "item_range_ratio"] <-  0.4705882
dat[dat$Item == "26:35:44", "item_range_ratio"] <-  0.5909091
dat[dat$Item == "36:45:54", "item_range_ratio"] <-  0.6666667

dat[dat$Item == "06:15:24", "item_range_ratio_scaled"] <-  -1.3441995
dat[dat$Item == "16:25:34", "item_range_ratio_scaled"] <-  -0.1316642
dat[dat$Item == "26:35:44", "item_range_ratio_scaled"] <-   0.5297187  
dat[dat$Item == "36:45:54", "item_range_ratio_scaled"] <-   0.9461450

dat[dat$Item == "06:15:24", "item_mean_ratio"] <-  0.5125000 
dat[dat$Item == "16:25:34", "item_mean_ratio"] <-  0.6876471 
dat[dat$Item == "26:35:44", "item_mean_ratio"] <-  0.7691558
dat[dat$Item == "36:45:54", "item_mean_ratio"] <-  0.8166667

dat[dat$Item == "06:15:24", "item_mean_ratio_scaled"] <-   -1.37582241 
dat[dat$Item == "16:25:34", "item_mean_ratio_scaled"] <-   -0.06614191 
dat[dat$Item == "26:35:44", "item_mean_ratio_scaled"] <-    0.54334858 
dat[dat$Item == "36:45:54", "item_mean_ratio_scaled"] <-    0.89861574

# VAGUENESS
# Create a factor coding for Vagueness
dat[ dat$Condition==1 , 'Vagueness'] <- 'Vague'
dat[ dat$Condition==2 , 'Vagueness'] <- 'Crisp'
dat[ dat$Condition==3 , 'Vagueness'] <- 'Vague'
dat[ dat$Condition==4 , 'Vagueness'] <- 'Crisp'
dat$Vagueness <- as.factor(dat$Vagueness)

# manually center Vagueness
dat$c_Vag <- ifelse(dat$Vagueness=='Crisp', -.5, .5)

dat$Selection = ""
dat[dat$Condition %in% c(1,2), "Selection"] <- "matching"
dat[dat$Condition %in% c(3,4), "Selection"] <- "comparison"
dat$Selection.sum <- ifelse(dat$Selection=="matching", .5, -.5)
dat$Selection <- as.factor(dat$Selection)

# remove one zero RT by replacing with NA
dat$RT[dat$RT==0]<-NA

# ORDER
# give the levels of Order meaningful names
dat$Order <- factor(dat$Order, levels=c(1,2), labels=c('LtoR','RtoL'))

# make a manually centred Order
dat$c_Ord <- ifelse(dat$Order=="LtoR",-.5,.5)

# QUANTITY
# give the levels of Quantity meaningful names
dat$Quantity <- factor(dat$Quantity, levels=c(1,2), labels=c('Small','Large'))

# make a manually centred Quantity
dat$c_Qty <- ifelse(dat$Quantity=="Small",-.5,.5)

# INSTRUCTION
# add number of characters in the instruction # 29 30 34 36 38
dat$nchar_instr = nchar(as.character(dat$Instruction))
dat$nchar_instr_scaled = as.vector(scale(nchar(as.character(dat$Instruction)), scale=TRUE))

# make Instruction be a factor (17 levels)
dat$Instruction <- as.factor(dat$Instruction)

# RT
# add transformations of RT
dat$RT_RecSqd    <- 1/dat$RT^2
dat$RT_Rec       <- 1/dat$RT
dat$RT_RecSqt    <- 1/sqrt(dat$RT) 
dat$RT_log       <- log(dat$RT)
dat$RT_sqt       <- sqrt(dat$RT)
dat$RT_raw       <- dat$RT
dat$RT_sqd       <- dat$RT^2

# save dat as dat.Rda
# This data set contains *all* trials including impossible trials and is mainly for graphs comparing different removals
save(dat, file="dat.Rda")

 # lose impossible trials
  dd <- dat
  dd$RT[dd$RT == 1 ] <- NA
  dd$RT[dd$RT > 40000 ] <- NA
  dd <- dd[complete.cases(dd),]
  row.names(dd) <- NULL

# add preceding RT: because we removed impossible trials, the value for preceding RT for a trial following an impossible trial is the value of the trial that preceded the impossible trial.
dd$RTprev <- NA
for (s in levels(dd$Subject)) {
  nrows = nrow(dd[dd$Subject==s,])
  for (i in 1:nrows) {
    if (i==1) {
      dd[dd$Subject==s, "RTprev"][i] <- dd[dd$Subject==s, "RT"][i]
    }
    else
      dd[dd$Subject==s, "RTprev"][i] <- dd[dd$Subject==s, "RT"][i-1]
  }
}

# add transformations of previous RT
dd$RTprev_RecSqd    <- 1/dd$RTprev^2
dd$RTprev_Rec       <- 1/dd$RTprev
dd$RTprev_RecSqt0   <- 1/sqrt(dd$RTprev) 
dd$RTprev_RecSqt    <- as.vector(scale( 1/sqrt(dd$RTprev)  ))
dd$RTprev_log       <- log(dd$RTprev)
dd$RTprev_sqt       <- sqrt(dd$RTprev)
dd$RTprev_raw       <- dd$RTprev
dd$RTprev_sqd       <- dd$RTprev^2

# save dd as dd.Rda
# This data set non-impossible trials and renamed/recomputed variables
save(dd, file="dd.Rda")
```

Check the distribution of the trimmed data to see if a transformation is suggested. Log transformation seems to be best so we adopt that for the rest of the analyses.

```{r boxplot subjects and items}
load("dd.Rda")
# allow reference to transformation
subdata = melt(dd,
               measure.vars=c("RT_Rec", "RT_RecSqt", "RT_log", "RT_raw"),
               variable.name="transformation",
               value.name="value") 
# boxplots for subjects and items separately for each dependent variable
ggplot(subdata) + 
  facet_wrap(~ variable, scales="free", ncol=4) +
  geom_boxplot(aes(y=value, x=Item, col="Items")) + 
  geom_boxplot(aes(y=value, x=Subject, col="Subjects")) + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1), aspect.ratio=1, panel.grid=element_blank()) + 
  scale_colour_manual("",values=c("Items"="red","Subjects"="blue")) +
  xlab("") + ylab("")
```

```{r load_dd}
load("dd.Rda")
```

```{r do_lmer}
v5 <- lmer(data=dd, 
           RT_RecSqt ~ 
             c_Vag + Selection.sum + c_Qty + c_Ord + 
             Selection.sum:c_Vag:c_Qty +
             item_mean_ratio + 
             s_Trl +
             RTprev_RecSqt +
             nchar_instr +
             (1+c_Vag + Selection.sum + c_Qty + c_Ord|Subject)
)
```

```{r summary_lmer}
summary(v5)
```

```{r R2}
cat("R^2")
cor(fitted(v5), dd$RT_RecSqt)^2
```

```{r plotLMER}
par(mfrow=c(2,5))
plotLMER.fnc(v5)
```

Plot model coefficients and ci's

```{r coeff}
par(las=2, mar=c(14,6,1,1))
y=as.data.frame(summary(v5)$coef[-1,1]) # estimates, without intercept
q=confint(v5)[-c(1,2,3),]
y=cbind(y,q)
plot(y[,1], ylim=range(y[,1],y[,2],y[,3]), xaxt="n", xlab="", ylab="estimate\n",  pch='-', cex=2)
segments(x0=1:nrow(y), x1=1:nrow(y), y[,2], y[,3])
abline(h=0)
axis(1, labels=row.names(y), at=1:nrow(y), srt=45)
grid()
```

Model criticism baayen plots

```{r baayenplots, fig.width=9, fig.height=3}
# Baayen 4-plot model criticism

par(mfrow=c(1,4), pty='s')
# create scaled residuals
v5$rstand = as.vector(scale(resid(v5)))
# plot scaled residuals density
plot(density(v5$rstand))
# plot sample quantiles versus theoretical quantiles
qqnorm(v5$rstand, cex=.5)
qqline(v5$rstand)
# plot standardised residuals versus fitted values
plot(v5$rstand ~ fitted(v5), pch='.')
# absolute standardised residuals greater than 2.5 are candidates for being outliers, the abline identifies them on the plot
abline(h=c(-2.5,2.5))
# create dffits
dffits=abs(resid(v5, "dffits"))
# plot dffits
plot(dffits, type='h')
```

>>>>

---
title: "Experiment 3 of 4"
author: "Matt Green"
output:
  html_document:
    toc: true
    <!--toc_float: true-->
    mathjax: default
    template: default
    <!--code_folding: show-->
    theme: default
---  
<style>
table, th, td {
    border: 1px grey;
    border-collapse: collapse;
    padding:2px 5px 2px 5px
}  
</style>

# Experiment 3

```{r loadLibraries, message=FALSE}
library(knitr); library(MASS); library(ggplot2); library(lme4); library(xtable); library(plyr); library(lattice); library(LMERConvenienceFunctions); library(car); library(grid); library(lmerTest); library(reshape2); library(ggplot2); library(languageR); library(pander); source('summarySEwithin2.R')
```

Experiment Structure: 
Each subject saw 192 trials, being 3 cycles of 64 trials. There were 38 participants. There were 7296 rows in the full data set.

Trial Procedure: 
First an instruction that was a referring expression was presented until participants dismissed it with a key press. Then the squares with dots were presented until the participant responded with a key press, with a 60 second timeout if they did not respond.

```{r showInstructions, echo=FALSE, cache=TRUE}
v = c("crisp", "crisp", "vague", "vague")
n = c('number', 'number', 'number', 'number')
t = c("matching", "comparison", "matching", "comparison")
e = c("Choose a square with 6 dots", "Choose a square with fewer than 20 dots", "Choose a square with about 10 dots", "Choose a square with far fewer than 20 dots")
i = data.frame(v, n, t, e)
names(i) = c("vagueness", "instruction format", "selection task", "example instruction")
pander(i, justify=c('center', 'left', 'left', 'left'))
```

```{r showItems, echo=FALSE, cache=TRUE}
items = c(1,2,3,4,5,6,7,8)
i = data.frame(Item = c("1","2","3","4"),
               x = c("06:15:24", "16:25:34", "26:35:44", "36:45:54"))
names(i) = c('Item', 'Numbers of dots')
pander(i, justify=c('center', 'center'))
```

```{r getData, echo=T}
rawdata <- read.table('data.txt', sep=' ')
```

```{r trimat25sd}
trimdata <- perSubjectTrim.fnc(rawdata, response='RT', subject='Subject', trim = 2.5)$data
rtdata <- trimdata
```
```{r computeSummariesForGraphs, cache=TRUE, message=FALSE}
rtdataplot1 <- summarySEwithin2(rtdata, measurevar="RT_log", withinvars=c("Condition", "Vagueness", "Selection", "Item"), idvar="Subject")
```

```{r rtplot, fig.width=7, fig.height=3, echo=F}
dodge = position_dodge(width=0.2)
rtdataplot1$Condition <- as.factor(paste(sep=' ', rtdataplot1$Selection, rtdataplot1$Vagueness))
rtplot=ggplot(rtdataplot1, aes(y=RT_logNormed, x=Item, ymin=RT_logNormed-ci, ymax=RT_logNormed+ci,
                group=Condition, shape=Condition, fill=Condition)) +
  geom_line(position=dodge) +
  geom_errorbar(width=.2, position=dodge) +
  geom_point(size=2, position=dodge) +
  scale_shape_manual("",values = c(22, 22, 21, 21)) +
  scale_fill_manual("",values=c("black","white","black","white")) +
  ggtitle("Response time") +
  ylab("Respone time (log (ms))") + 
  xlab("") +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        legend.key = element_blank(), aspect.ratio=1,
        axis.text.x = element_text(angle = 45, hjust = 1),
        plot.background=element_rect(fill=NA, color='white')) +
  facet_grid(~Selection) 
rtplot
```

```{r rtlmer, cache=TRUE}
rtlmer = lmerTest::lmer(data=rtdata,
                        RT_log ~ c_Vag * c_Sel + c_Itm +
                          (1 + c_Vag * c_Sel + c_Itm | Subject))
```

```{r rtlmersummaryxtable, results='asis', cache=TRUE, echo=F}
print(xtable(summary(rtlmer)$coef, digits=c(0,2,2,1,1,4)),type='html')
```

```{r seplmersforcomparison, results='asis', cache=TRUE, echo=-3}
s1 <- subset(rtdata, subset=Selection=='Comparison')
s1lm <- lmerTest::lmer(data=s1, RT_log ~ c_Vag + c_Itm + (1 + c_Vag + c_Itm | Subject))
print(xtable(summary(s1lm)$coef, digits=c(0,2,2,1,1,4)), type='html')
```

```{r seplmersformatching, results='asis', cache=TRUE, echo=-3}
s2 <- subset(rtdata, subset=Selection=='Matching')
s2lm <- lmerTest::lmer(data=s2, RT_log ~ c_Vag + c_Itm + (1 + c_Vag + c_Itm | Subject))
print(xtable(summary(s2lm)$coef, digits=c(0,2,2,1,1,4)), type='html')
```

### 2016-09-27 experiments/e3/writeup/preprocessing/preprocessing.Rmd
e3 writeup less verbose

---
title: "Experiment 3 of 4"
output:
  html_document:
    toc: true
    <!--toc_float: true-->
    mathjax: default
    template: default
    code_folding: hide
    theme: default
---  
# Experiment 3

```{r docOpts}
library(knitr)
```

```{r loadLibraries}
suppressPackageStartupMessages(library(MASS))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(lme4))
suppressPackageStartupMessages(library(xtable))
suppressPackageStartupMessages(library(plyr))
suppressPackageStartupMessages(library(reshape))
suppressPackageStartupMessages(library(lattice))
suppressPackageStartupMessages(library(LMERConvenienceFunctions))
suppressPackageStartupMessages(library(car))
suppressPackageStartupMessages(library(grid))
suppressPackageStartupMessages(library(lmerTest))
suppressPackageStartupMessages(library(reshape2))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(require(languageR))
```

# Experiment Structure
Each subject saw 192 trials, being 3 cycles of 64 trials. There were 38 participants. There were 7296 rows in the full data set.

# Trial Procedure
The procedure for each trial was as follows. First an instruction that was a referring expression was presented until participants dismissed it with a key press. Then the squares with dots were presented until the participant responded with a key press, with a 60 second timeout if they did not respond.

# Prepare the data
Source additional R files with functions for computing confidence intervals.

```{r sourceFunctions}
source("functions/CIfunctions.R")
```

Get data.

```{r readdata}
dat <- read.delim("data/allresults.txt", stringsAsFactors = FALSE)

@@

# sort out borderline responses into expected near and far
# How many squares in the square they chose? 
dat$RESPONSE <- as.character(dat$RESPONSE)
for (row in 1 : nrow(dat) ) {
  switch(dat[row,'RESPONSE'],
         'LEFT' = {dat[row, 'choice'] <- dat[row, 'Left']},
         'MIDDLE' = {dat[row, 'choice'] <- dat[row, 'Mid']},
         'RIGHT' = {dat[row, 'choice'] <- dat[row, 'Right']}         
  )
}
dat$crossed=paste('Con',dat$Condition,':Quan',dat$Quantity,':Item',dat$Item,sep="")

@@

dat$Selection = ""
dat[dat$Condition %in% c(1,2), "Selection"] <- "matching"
dat[dat$Condition %in% c(3,4), "Selection"] <- "comparison"
dat$Selection.sum <- ifelse(dat$Selection=="matching", .5, -.5)
dat$Selection <- as.factor(dat$Selection)

@@

# save dd as dd.Rda
# This data set non-impossible trials and renamed/recomputed variables
save(dd, file="dd.Rda")
```

Check the distribution of the trimmed data to see if a transformation is suggested. Log transformation seems to be best so we adopt that for the rest of the analyses.

```{r boxplot subjects and items}
load("dd.Rda")
# allow reference to transformation
subdata = melt(dd,
               measure.vars=c("RT_Rec", "RT_RecSqt", "RT_log", "RT_raw"),
               variable.name="transformation",
               value.name="value") 
# boxplots for subjects and items separately for each dependent variable
ggplot(subdata) + 
  facet_wrap(~ variable, scales="free", ncol=4) +
  geom_boxplot(aes(y=value, x=Item, col="Items")) + 
  geom_boxplot(aes(y=value, x=Subject, col="Subjects")) + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1), aspect.ratio=1, panel.grid=element_blank()) + 
  scale_colour_manual("",values=c("Items"="red","Subjects"="blue")) +
  xlab("") + ylab("")
```

```{r load_dd}
load("dd.Rda")
```

```{r do_lmer}
v5 <- lmer(data=dd, 
           RT_RecSqt ~ 
             c_Vag + Selection.sum + c_Qty + c_Ord + 
             Selection.sum:c_Vag:c_Qty +
             item_mean_ratio + 
             s_Trl +
             RTprev_RecSqt +
             nchar_instr +
             (1+c_Vag + Selection.sum + c_Qty + c_Ord|Subject)
)
```

```{r summary_lmer}
summary(v5)
```

```{r R2}
cat("R^2")
cor(fitted(v5), dd$RT_RecSqt)^2
```

```{r plotLMER}
par(mfrow=c(2,5))
plotLMER.fnc(v5)
```

Plot model coefficients and ci's

```{r coeff}
par(las=2, mar=c(14,6,1,1))
y=as.data.frame(summary(v5)$coef[-1,1]) # estimates, without intercept
q=confint(v5)[-c(1,2,3),]
y=cbind(y,q)
plot(y[,1], ylim=range(y[,1],y[,2],y[,3]), xaxt="n", xlab="", ylab="estimate\n",  pch='-', cex=2)
segments(x0=1:nrow(y), x1=1:nrow(y), y[,2], y[,3])
abline(h=0)
axis(1, labels=row.names(y), at=1:nrow(y), srt=45)
grid()
```

Model criticism baayen plots

```{r baayenplots, fig.width=9, fig.height=3}
# Baayen 4-plot model criticism

par(mfrow=c(1,4), pty='s')
# create scaled residuals
v5$rstand = as.vector(scale(resid(v5)))
# plot scaled residuals density
plot(density(v5$rstand))
# plot sample quantiles versus theoretical quantiles
qqnorm(v5$rstand, cex=.5)
qqline(v5$rstand)
# plot standardised residuals versus fitted values
plot(v5$rstand ~ fitted(v5), pch='.')
# absolute standardised residuals greater than 2.5 are candidates for being outliers, the abline identifies them on the plot
abline(h=c(-2.5,2.5))
# create dffits
dffits=abs(resid(v5, "dffits"))
# plot dffits
plot(dffits, type='h')
```

>>>>

```{r readdata}
dat <- read.delim("../rawdata.txt", stringsAsFactors = FALSE)

@@

# sort out borderline responses into expected near and far
dat$RESPONSE <- as.character(dat$RESPONSE)
for (row in 1 : nrow(dat) ) {
  switch(dat[row,'RESPONSE'],
         'LEFT' = {dat[row, 'choice'] <- dat[row, 'Left']},
         'MIDDLE' = {dat[row, 'choice'] <- dat[row, 'Mid']},
         'RIGHT' = {dat[row, 'choice'] <- dat[row, 'Right']}         
  )
}
dat$crossed=paste('Con',dat$Condition,':Quan',dat$Quantity,':Item',dat$Item,sep="")

@@

dat$Selection = ""
dat[dat$Condition %in% c(3,4), "Selection"] <- "Comparison"
dat[dat$Condition %in% c(1,2), "Selection"] <- "Matching"
dat$c_Sel <- ifelse(dat$Selection=="Comparison", -.5, .5)
dat$Selection <- as.factor(dat$Selection)

@@

# save dd as data.Rda
# This data set non-impossible trials and renamed/recomputed variables
write.table(dd, file="../data.txt")
```
